---
title: '5 State ABM Tutorial'
subtitle: ''
author: "Jasper Zhang"
output:
  html_document: default
  pdf_document: default
---


-   Jasper Zhang, MSc
-   Petros Pechlivanoglou, PhD


Please acknowledge our work. See details to cite below.

\newpage

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE, eval = TRUE)
options(scipen = 999)  # disable scientific notation
rm(list = ls())        # clear memory (removes all the variables from the workspace)
```

Change `eval` to `TRUE` if you want to knit this document.

# 01 Load packages

```{r}
if (!require('pacman')) install.packages('pacman'); library(pacman) 
# load (install if required) packages from CRAN
p_load("devtools", "dplyr", "scales", "ellipse", "ggplot2", "lazyeval", "igraph", "truncnorm", "ggraph", "reshape2", "knitr", "markdown", "stringr", "dampack", "matrixStats")
# load (install if required) packages from GitHub
# install_github("DARTH-git/darthtools", force = TRUE) # Uncomment if there is a newer version
p_load_gh("DARTH-git/darthtools")
```

# 02 Load functions 

```{r}
#negate function
`%!in%`                      <- Negate(`%in%`)
```


In this basic ABM we assume that individuals can be be in one of 5 states:  1. susceptible to an infection, 2.infected and infectious but not in quarantine, 3.infected in quarantine, 4. recovered. Recovery is assumed to be a saturating state. We assumed that the quarantine period is 14 days from infection for all infected individuals. We further assumed that for the first day of the disease individuals are infectious although asymptomatic, and that as soon as there are symptoms individuals will self-quarantine. We further assumed perfect adherence to quarantine protocols. 
We developed a agent based model where we assumed a daily cycle length. We modelled a hypothetical environment where indiciduals will go to work and can be infected at work and in the community but cannot infect others at home. We used the model to prospectively estimate a 60 day period. We assumed a latency of 4 days from acquiring the infection until symptom development 



We used seeding to ensure that the results were reproducible

# 03 Model input

## 03.1 Define model input parameters

```{r}
## General setup
set.seed(1)                            # set the seed 
cycle_length    <- 1                   # cycle length equal to one day
n_cycles        <- 60                  # number of cycles
n_i             <- 100#00              # number of individuals
q_maxtime       <- 14                  # quarantine period
latentperiod    <- 4                   # time from exposure to becoming infectious


# the 5 health states of the model:
v_names_states  <- c("S",   # Susceptble 
                     "SQ",  # Susceptble quarantine 
                     "I",   # Infected
                     "IQ",  # Infected quarantine
                     "R")   # Recovered      

v_names_cycles  <- paste("cycle", 0:n_cycles)    # cycle names
n_states        <- length(v_names_states)        # number of health states 

#PP: not sure this is necessary

#define groups of states

ending_states                <- c("R")
infect_states                <- c("I","IQ")
infect_states_iso            <- c("IQ")
all_infect_states            <- c(infect_states,infect_states_iso)


iso_states                   <- c("IR","IQ")

sus_states                   <- c("S","SQ")


```
We assume a 


```{r}
#### Deterministic analysis ####

# input for Creating the synthetic cohort'
n_S <- round(0.7 * npop) # number of susceptible at time 0



#Transition probabilities
#(all terminal probabilities are conditional on not being admitted/recovered)

#probability of being detected
p_detect_overall <- 0.3
p_detect_inf <- 0.6

#infection probabilities vary by location
#household
p_transhh_young_child <- 0.0160176427409341  #  <= 10 
p_transhh_old_child   <- 0.077154345991161   #  <=17  
p_transhh_adult       <- 0.076831649685427   #  > 18

#vector of household transmission probabilites
p_transhh = c(p_transhh_young_child, p_transhh_old_child,  p_transhh_adult)

#atwork \ university
p_TransmitWork        <- 0.0220595686685414




```


# 04 Sample individual level characteristics

## 04.1 Static characteristics

```{r}
#initialize population dataframe
create_abmpop <- function(npop){
  # Arguments:
  # npop: number of people in the population
  # Returns: 
  # new population dataframe with initialized states
  
  
  #create matrix of agent
  agents = data.frame(ptid = 1:npop, # assign id 
                      Sex = sample(x = c("Female", "Male"),
                                   prob = c(0.5, 0.5), 
                                   size = npop, 
                                   replace = TRUE) , #Randomly generate gender
                      Age = round(rnorm(npop, mean=30, sd=15), 
                                  digits = 0), # draw Age from a normal distribtion
                      wid = 0, #workplace id
                      fid = 0, #family id
                      qtime = 0, #quarantine time clock
                      time = 0, # number of total quarantine times
                      state = "S") #initialized state of everyone is healthy
  
  #intial state
  
  # Number of 'S' and 'I' based on 7:3 ratio
  
  n_I <- npop - n_S
  # Create the initial state assignments
  agents$state <- sample(c(rep("S", n_S), rep("I", n_I)))
  
# Add age_group based on Age
  agents$Age[agents$Age < 19] <- 19 # only adults
  agents$Age_group <- ifelse(agents$Age <= 10, 1,
                         ifelse(agents$Age <= 17, 2, 3))
  
  family_size = c(1,2,3,4) # family size can be 1 2 3 4
  p_family_size = family_size/sum(family_size) # weights of each family size
  
  fam_num = floor(npop * p_family_size / family_size)[1]
  fam_id  = c(rep(1:fam_num,each = family_size[1]),
              rep((fam_num + 1): (fam_num*2),each=family_size[2]),
              rep((fam_num*2 + 1):(fam_num*3),each=family_size[3]),
              rep((fam_num*3 + 1):(fam_num*4),each=family_size[4]))
  
  fam_id = c(fam_id,rep((max(fam_id) + 1),each = (npop - length(fam_id))))
  
  work_size = c(5,7,8) # workplace size is set to 5,7,8
  p_work_size = work_size/sum(work_size) # weights of each work size
  work_num = floor(npop * p_work_size / work_size)[1]
  work_id  = c(rep(1:work_num,each = work_size[1]),
               rep((work_num + 1): (work_num*2),each=work_size[2]),
               rep((work_num*2 + 1):(work_num*3),each=work_size[3]))
  
  work_id = sample(c(work_id,rep((max(work_id) + 1),each = (npop - length(work_id)))))
  
  agents$fid = fam_id
  agents$wid = work_id
  
  
  agents$expl           <- 0 # exposure level: number of infected people contacted within a day
  agents$expl_home      <- 0 # exposure level: number of infected people contacted at home within a day
  agents$contact_num    <- 0 # number of contacts a day
  agents$contact_max    <- 0 # max number of contact
  agents$contact_type   <- 0 # contact type
  agents$atwork         <- 1
  agents$v_Ts           <- 0     # time of being infected
  agents$detect_time    <- 5
  agents$new_detection  <- 0
 
  
  #set number of contacts atwork
  work_in_sample             <- agents$ptid[which(agents$atwork == TRUE)]
  work_size                  <- table(agents[work_in_sample,]$wid) #workplacesize
  work_ppl                   <- agents[work_in_sample,]
  work_info                  <- data.frame(wid = as.numeric(as.character(names(work_size))),wsize =  as.numeric(as.matrix(work_size)))
  work_place_size            <- merge(work_ppl, work_info, by = "wid")
  work_place_size            <- work_place_size[order(work_place_size$ptid),]
  
  work_place_size <- work_place_size %>%
    rowwise() %>%
    mutate(contact_num = sample(0:(wsize-1), 1)) # randomly generate contact number from 0 to worksize -1

  agents$contact_num[work_in_sample]  <- work_place_size$contact_num
  agents$contact_max[work_in_sample]  <- work_place_size$wsize -1
  agents$contact_type[work_in_sample] <- 1 #atworks
  
  
  return(agents)

}





```

## 04.2 Dynamic characteristics

```{r}
# Specify the initial health state of the individuals 
# everyone begins in the healthy state (in this example)
v_M_init  <- rep("S", times = n_i)   
v_n_cycles_s_init <- rep(0, n_i)  # a vector with the time of being sick at the start of the model 
```

## 04.3 Create a dataframe with the individual characteristics

```{r}
# create a data frame with each individual's 
# ID number, treatment effect modifier, age and initial time in sick state 

# NOTE: we use n_cycles_s for the number of times being sick, we start the data frame with the initial "history" of time being sick by saving v_Ts_init. However, during the simulation this value is updated and therefore called numer of times being sick.
#head(df_X) # print the first rows of the dataframe

df_X <- create_abmpop(n_i)

```

# 05 Define Simulation Functions

## 05.1 Probability function

The `Probs` function updates the transition probabilities of every cycle is shown below.

```{r}

infect_prob <- function(prob,exposure) { 
  # Arguments:
  # prob:     the probability of infection in one time contact
  # exposure: the time of being exposed to infectious individuals
  # Returns: 
  # the probability of being infected based on certain exposure time
  return(1 - (1 - prob) ^ (exposure))   # 1 - prob is the prob of being safe, 
  # 1 minus the prob of always being safe is the prob of getting infection
}

Probs <- function(M_t, df_X, v_Ts) { 
  # Arguments:
  # M_t:  health state occupied at cycle t (character variable)
  # df_X: data frame with individual characteristics data 
  # v_Ts: vector with the duration of being infected
  # Returns: 
  # transition probabilities for that cycle
  
  # create matrix of state transition probabilities
  m_p_t           <- matrix(0, nrow = n_states, ncol = n_i)  
  # give the state names to the rows
  rownames(m_p_t) <-  v_names_states                            
  
  #Calculate the probability of being infected for susceptible people
  sus <- df_X[df_X$state == "S",] #identify those susceptible
  
  #check exposure number
  sus_prob_home  <- rep(0, nrow(sus)) 
  contact_active <- sus[sus$expl > 0 | sus$expl_home > 0, ]
  at_home        <- contact_active$expl_home > 0
  
  if(sum(at_home) > 0){
    home_infect = infect_prob(p_transhh[contact_active[at_home,]$age_group],contact_active[at_home,]$expl_home)
    sus_prob_home[sus$expl_home > 0] = home_infect
  }
  
  
  sus_prob = rep(0,nrow(sus))
  at_work = contact_active$contact_type == 1
  if(sum(at_work) > 0){
    work_infect = infect_prob(p_TransmitWork,contact_active[at_work,]$expl)
    sus_prob[sus$expl > 0 & sus$contact_type == 1] = work_infect
  }
  p_Sus_Inf <- (1- ((1- sus_prob_home) * (1 - sus_prob))) # 1 minus the prob of being safe all day
  

  #Calculate the probability of being infected for susceptible quarantine
  sus_Q <- df_X[df_X$state == "SQ",] #identify those susceptible
  
  #check exposure number
  sus_prob_home_Q  <- rep(0, nrow(sus_Q)) 
  contact_active_Q <- sus[sus_Q$expl > 0 | sus_Q$expl_home > 0, ]
  at_home_Q        <- contact_active$expl_home > 0
  
  if(sum(at_home_Q) > 0){
    home_infect_Q = infect_prob(p_transhh[contact_active[at_home_Q,]$age_group],contact_active_Q[at_home_Q,]$expl_home)
    sus_prob_home_Q[sus_Q$expl_home > 0] = home_infect_Q
  }

  
  
  
  # the 5 health states of the model:
v_names_states  <- c("S",   # Susceptble 
                     "SQ",  # Susceptble quarantine 
                     "I",   # Infected
                     "IQ",  # Infected quarantine
                     "R")   # Recovered      

  
  #susceptible 
  m_p_t["S"                                   , M_t == "S"]                    <- 1 - p_Sus_Inf
  m_p_t["SQ"                                  , M_t == "S"]                    <- 0
  m_p_t["I"                                   , M_t == "S"]                    <- p_Sus_Inf
  m_p_t["IQ"                                  , M_t == "S"]                    <- 0
  m_p_t["R"                                   , M_t == "S"]                    <- 0
  
  
  #susceptible quarantine
  m_p_t["S"                                   , M_t == "SQ"]                   <- 1 -  sus_prob_home_Q
  m_p_t["SQ"                                  , M_t == "SQ"]                   <- 0
  m_p_t["I"                                   , M_t == "SQ"]                   <- 0
  m_p_t["IQ"                                  , M_t == "SQ"]                   <- sus_prob_home_Q
  m_p_t["R"                                   , M_t == "SQ"]                   <- 0
  
  #infected
  m_p_t["S"                                   , M_t == "I" ]                   <- 0
  m_p_t["SQ"                                  , M_t == "I" ]                   <- 0
  m_p_t["I"                                   , M_t == "I"]                    <- 1
  m_p_t["IQ"                                  , M_t == "I"]                    <- 0
  m_p_t["R"                                   , M_t == "I"]                    <- 0
  
  #infected quarantine
  m_p_t["S"                                   , M_t == "IQ"]                   <- 0
  m_p_t["SQ"                                  , M_t == "IQ"]                   <- 0
  m_p_t["I"                                   , M_t == "IQ"]                   <- 0
  m_p_t["IQ"                                  , M_t == "IQ"]                   <- 1
  m_p_t["R"                                   , M_t == "IQ"]                   <- 0
  
  
  #recover ending state
  m_p_t["S"                                   , M_t == "R"]                    <- 0
  m_p_t["SQ"                                  , M_t == "R"]                    <- 0
  m_p_t["I"                                   , M_t == "R"]                    <- 0
  m_p_t["IQ"                                  , M_t == "R"]                    <- 0
  m_p_t["R"                                   , M_t == "R"]                    <- 1
   
  
  return(t(m_p_t))
}  


```

## 05.1 Quanrantine Function

```{r}

#Quarantine_Check can update the quarantine clock at current time cycle
Quarantine_Check <- function(pop, pre_state){
  # Arguments:
  # pop: population dataframe of current time cycle t
  # pre_state: states at previously cycle t-1
  # Returns: 
  # new population dataframe with updated states
  
  # Step1: find people in different quarantine conditions
  
  # People in quarantine:
  # condition 1: reset quarantine clock for those who become infected or have close contact with infected people during quarantine
  new_infected_in_Q          = which(pop$state %in% infect_states_iso & pre_state == "SQ") #got infected in quarantine
  
  reset_family               = unique(pop$fid[new_infected_in_Q])
  reset_people               = pop$fid %in% reset_family #find all the people who had expose in family quarantine
  
  
  # condition 2: find people are going to be released from quarantine
  # check if quarantine clock hit q_maxtime
  full_q                     = pop$qtime == q_maxtime & !reset_people & pop$state %!in% ending_states 
  
  #find susceptible people and infected people who meet quarantine time
  sus_full_q                 = full_q & pop$state == "SQ"
  inf_full_q                 = full_q & pop$state %in% infect_states_iso 
  
  # condition 3: find people who stay in quarantine and the quarantine clock has to be incremented.
  else_q                     = (pop$qtime < q_maxtime & pop$qtime >0) & !reset_people & pop$state %!in% ending_states  
  
  # Other people:
  # condition 4: find newly infected people entering quarantine NOTE: people enter quarantine by detection
  new_infected               = which(pop$state %in% iso_states & pop$qtime == 0) #new infected generated from infected
  
  # condition q_maxtime: find newly infected people's dependencies in families and workpalces
  new_infected_family        = unique(pop$fid[new_infected]) #families have newly infected in quarantine cases
  
  #new_infected_work= unique(pop$wid[new_infected]) #workplaces have newly infected in quarantine cases #comment out if same workplace contact don't need to be put in quarantine
  
  new_qurantine              = (((pop$fid %in% new_infected_family & pop$qtime == 0) | #(pop$wid %in% new_infected_work & pop$qtime == 0) |
                                   (pop$fid %in% reset_family)) & (pop$state %!in% ending_states) & (!(pop$essential == 1 & pop$getspsl == 0))  ) #people whose quarantine clocks have to be set to 1 including
  new_qurantine_add = new_qurantine & pop$state %!in% iso_states #find people need to be put in quarantine state
  
  
  # Step2: Update quarantine clock
  # set new_quarantine clock to 1 
  pop$qtime[new_qurantine]    = 1
  pop$state[new_qurantine_add] = paste(pop$state[new_qurantine_add],"Q",sep = "") #update quarantine states
  pop$time[new_qurantine]     = pop$time[new_qurantine]+1 #increment total quarantine time
  
  
  # release people with meet q_maxtime-days quarantine requirement
  # - isolated susceptible people will go to susceptible
  # - infected people will go to recovered
  pop$qtime[full_q]           = 0
  pop$state[sus_full_q]       = "S" #when quarantine ends, state goes back to susceptible
  pop$state[inf_full_q]       = "R" #when quarantine ends, state goes back to recovered
  
  #increment quarantine clock and total quarantine time for everyone else in quarantine
  pop$qtime[else_q]           = pop$qtime[else_q] + 1
  pop$time[else_q]            = pop$time[else_q] + 1
  return (pop)
}

```

## 0.5.1 Exposure at work

```{r}

#Exposure_Check Work
work_exposure <- function(pop){
  allpop                      = pop
  worker_list = pop[pop$atwork == 1,]
  
  if(sum(pop$atwork == 1) > 0){
    select_pop                = allpop[pop$atwork == 1,]
    expl_list                 = select_pop$expl
    infectious                = select_pop$state %in% infect_states # defn of infectious
    wid_list                  = select_pop$wid
    out_of_iso                = select_pop$state %!in% iso_states
    ptid_list                 = select_pop$ptid[out_of_iso & select_pop$state == "S"]
    atwork                    = select_pop$atwork
    
    workplacetable                = data.frame(table(wid_list[atwork]))
    
    names(workplacetable)         = c("wid","wsize")
    
    inf_worker               = select_pop[infectious ,]
    inf_workplace                = inf_worker$wid
    
    if (nrow(inf_worker) > 0){
      contact                   <- c()
      
      for( i in c(1:nrow(inf_worker))){
        select_worker           = inf_worker[i,]
        workplaceptid           = select_pop[select_pop$wid == select_worker$wid,]$ptid
        to_contact              = workplaceptid[workplaceptid %!in% select_worker$ptid]
        if (length(to_contact) > 1){
          new_contact = sample(to_contact,
                               min(length(to_contact),select_worker$contact_num))
          contact                 = c(contact,new_contact)
        }else if (length(to_contact) == 1){
          new_contact = to_contact
          contact                 = c(contact,new_contact)
        }
      }
      
      if (length(contact) != 0){
        contact_table             <- data.frame(table(contact))
        names(contact_table)      = c("ptid","expl")
        
        contact_table$ptid        <- as.numeric(as.character(contact_table$ptid))
        contact_table$expl        <- as.numeric(as.character(contact_table$expl))
        
        expl_list[select_pop$ptid %in% contact_table$ptid] = expl_list[select_pop$ptid %in% contact_table$ptid] + contact_table$expl
        allpop[(allpop$ptid %in% contact_table$ptid ),]$contact_type = 3
      }
    
    }
    allpop[pop$atwork == 1,]$expl  <- expl_list
    
  }
  
  return(allpop) 
}




```

## 05.2 Detection function

The `Detection` function detec infected people

```{r}
Detection <- function(pop,v_Ts){
  detection_period = v_Ts - latentperiod
  #detection_period[pop$sympt == "S"] = detection_period[pop$sympt == "S"] - incubation# individual won't get detection before develop symptoms.
  detected_individuals = (detection_period == pop$detect_time) & pop$state %!in% ending_states
  pop$new_detection = sum(detected_individuals)
  new_iso = detected_individuals & pop$state %!in% iso_states
  pop$state[new_iso ] = paste(pop$state[new_iso ],"Q",sep = "")
  return(pop)
  }


```


## 05.4 Microsimulation function

Below we develop the microsimulation function that allows the model to be run.

```{r}
MicroSim <- function(n_i = 100, df_X, seed = 1, n_cycles = 10) {
  # Arguments:  
    # n_i: number of individuals
    # df_X: data frame with individual data 
    # Trt: treatment
    # seed: seed for the random number generator, default is 1
    # cycle_length: cycle lenght 
  # Returns:
    # results: data frame with total cost and QALYs
  
  set.seed(seed) # set a seed to be able to reproduce the same results
  
  # create three matrices called m_M, m_C and m_E
  # number of rows is equal to the n_i, the number of columns is equal to n_cycles 
  # (the initial state and all the n_cycles cycles)
  # m_M is used to store the health state information over time for every individual
  # m_C is used to store the costs information over time for every individual
  # m_E is used to store the effects information over time for every individual
  
  m_M <-  matrix(nrow = n_i, ncol = n_cycles + 1, 
                               dimnames = list(paste("ind"  , 1:n_i, sep = " "), 
                                               paste("cycle", 0:n_cycles, sep = " ")))  
 
  m_vts    <-  matrix(nrow = n_i, ncol = n_cycles + 1, #tracking number infected days
                    dimnames = list(paste("ind"  , 1:n_i, sep = " "), 
                                    paste("cycle", 0:n_cycles, sep = " ")))   
  
  m_expl <-  matrix(nrow = n_i, ncol = n_cycles, 
                 dimnames = list(paste("ind"  , 1:n_i, sep = " "), 
                                 paste("cycle", 1:n_cycles, sep = " ")))  

  m_M[, 1] <- as.character(df_X$state) # initial health state at cycle 0 for individual i
  m_vts[,1]        <- df_X$v_Ts  
  v_Ts             <- df_X$v_Ts      # initialize time since illness onset
  v_inf            <- rep(0, n_i)    # initialize time of being infected


  v_total_inf_daily <- rep(0,n_cycles)    # initialize the number of total infections per cycle
  v_new_inf_daily   <- rep(0,n_cycles)    # initialize the number of total new infections per day
  v_detected_daily  <- rep(0,n_cycles)    # initialize the number of total infections per cycle
  
  
  # open a loop for time running cycles 1 to n_cycles 
  for (t in 1:n_cycles) {
     print(paste("running day: ",as.character(t),sep = ""))
    
    #check exposure
    df_X$expl            <- 0                                 # reset expl level
    df_X$expl_home       <- 0                                 # reset expl level
    df_X$contact_type    <- 0                                 # reset contact type
    df_X <- work_exposure(df_X)
    m_expl[,t] <- df_X$expl
    
    
    # calculate the transition probabilities for the cycle based on health state t
    m_P <- Probs(m_M[, t], df_X, v_Ts)
    # check if transition probabilities are between 0 and 1
    check_transition_probability(m_P, verbose = TRUE)
    # check if each of the rows of the transition probabilities matrix sum to one
    check_sum_of_transition_array(m_P, n_rows = n_i, n_cycles = n_cycles, verbose = TRUE)
    
    
    prob_new_state = samplev(m_P) #generate new states based on the prob
    df_X$state = prob_new_state
    df_X = Detection(df_X,v_Ts)
    prob_new_state = df_X$state
    
    v_detected_daily[t] = df_X$new_detection[1]
    v_total_inf_daily[t] = sum(prob_new_state %in% all_infect_states)
    v_new_inf_daily[t] = sum((m_M[, t] %in% sus_states   & prob_new_state %in% all_infect_states ))
    
    
    #update Quarantine Status
    df_X = Quarantine_Check(df_X,m_M[, t]) #m_M[,t] is the previous day's state
    qc_state = df_X$state # state_after quarantine check
    
    # update time since illness onset for t + 1 
    v_Ts <- if_else(qc_state %in% all_infect_states, v_Ts + 1, 0)
    m_vts[,t + 1] =  v_Ts 
    
    v_go_to_recover <- if_else(v_Ts >= 17, TRUE, FALSE)
    df_X$state[v_go_to_recover] = "R"
    
    m_M[, t + 1]  <- df_X$state  # new day's state
   
    
    # Display simulation progress
    if(t/(n_cycles/10) == round(t/(n_cycles/10), 0)) { # display progress every 10%
      cat('\r', paste(t/n_cycles * 100, "% done", sep = " "))
    }
    
  } # close the loop for the time points 
  
  # store the results from the simulation in a list
  results <- list(m_M = m_M ,
                  pop = df_X, 
                  v_new_inf_daily = v_new_inf_daily, 
                  v_total_inf_daily =   v_total_inf_daily ,
                  v_detected_daily =  v_detected_daily)   
  
  return(results)  # return the results

} # end of the `MicroSim` function  
```

## 06 Run Microsimulation

```{r, eval = TRUE, echo = TRUE, warning = FALSE, message = FALSE, results = FALSE }
# 06 Run Microsimulation

outcomes <- MicroSim(n_i = n_i, df_X = df_X, seed = 1)

# By specifying all the arguments in the `MicroSim()` the simulation can be started
# Run the simulation model
#outcomes_SoC  <- MicroSim(n_i = n_i, df_X = df_X, seed = 1, cycle_length = cycle_length, Trt = "SoC")
#$outcomes_trtA <- MicroSim(n_i = n_i, df_X = df_X, seed = 1, cycle_length = cycle_length, Trt = "A")
#outcomes_trtB <- MicroSim(n_i = n_i, df_X = df_X, seed = 1, cycle_length = cycle_length, Trt = "B")
```

# 07 Visualize results to be updated

```{r}
# Standard of Care
plot(density(outcomes_SoC$tc), main = paste("Total cost per person"),  xlab = "Cost ($)")
plot(density(outcomes_SoC$te), main = paste("Total QALYs per person"), xlab = "QALYs")
plot_trace_microsim(outcomes_SoC$m_M)      # health state trace
```

```{r}
# Treatment A
plot(density(outcomes_trtA$tc), main = paste("Total cost per person"),  xlab = "Cost ($)")
plot(density(outcomes_trtA$te), main = paste("Total QALYs per person"), xlab = "QALYs")
plot_trace_microsim(outcomes_trtA$m_M)     # health state trace
```

```{r}
# Treatment B
plot(density(outcomes_trtB$tc), main = paste("Total cost per person"),  xlab = "Cost ($)")
plot(density(outcomes_trtB$te), main = paste("Total QALYs per person"), xlab = "QALYs")
plot_trace_microsim(outcomes_trtB$m_M)     # health state trace
```

# 08 Cost-effectiveness analysis (CEA)

```{r}
# store the mean costs of each strategy in a new variable C (vector of costs)
v_C <- c(outcomes_SoC$tc_hat, outcomes_trtA$tc_hat, outcomes_trtB$tc_hat)
# store the mean QALYs of each strategy in a new variable E (vector of effects)
v_E <- c(outcomes_SoC$te_hat, outcomes_trtA$te_hat, outcomes_trtB$te_hat)

# use dampack to calculate the ICER
df_cea <- calculate_icers(cost       = v_C,
                          effect     = v_E,
                          strategies = v_names_str)
df_cea
```

```{R}
## CEA table in proper format 
table_cea <- format_table_cea(df_cea) 
table_cea
```

```{r}
## CEA frontier 
plot(df_cea, label = "all", txtsize = 14) +
  expand_limits(x = max(table_cea$QALYs) + 0.1) +
  theme(legend.position = c(0.8, 0.3))
```

We kindly request you to add the following Acknowledgement paragraph to your further work where DARTH code formed the basis. We also like to remind you that you can add other sources of reference to this paragraph to acknowledge code you got from others.

# Acknowlegdement

For this work we made use of the template developed by the Decision Analysis in R for Technologies in Health (DARTH) workgroup: <http://darthworkgroup.com>.

The notation of our code is based on the following provided framework and coding convention: Alarid-Escudero, F., Krijkamp, E., Pechlivanoglou, P. et al. A Need for Change! A Coding Framework for Improving Transparency in Decision Modeling. PharmacoEconomics 37, 1329--1339 (2019). <https://doi.org/10.1007/s40273-019-00837-x>.

Other work from DARTH can be found on the website: <http://darthworkgroup.com/publications/>

# Copyright for assignment work

Copyright 2017, THE HOSPITAL FOR SICK CHILDREN AND THE COLLABORATING INSTITUTIONS.All rights reserved in Canada, the United States and worldwide. Copyright, trademarks, trade names and any and all associated intellectual property are exclusively owned by THE HOSPITAL FOR Sick CHILDREN and the collaborating institutions. These materials may be used, reproduced, modified, distributed and adapted with proper attribution.
