---
title: '5-State Infectious Disease Microsimulation'
subtitle: 'Agent-based model with quarantine, workplace and household transmission'
author: "Jasper Zhang"
output:
  html_document: default
  pdf_document: default
---

**Authors:** Jasper Zhang (MSc), Petros Pechlivanoglou (PhD)

\newpage

## Model statistics and key parameters

| Parameter | Value | Description |
|-----------|-------|-------------|
| **Individuals** | 1,000 | Population size (`n_i`) |
| **Health states** | 5 | S, SQ, I, IQ, R |
| **Simulation horizon** | 60 days | Number of daily cycles (`n_cycles`) |
| **Quarantine duration** | 14 days | Required quarantine length (`q_maxtime`) |
| **Detection day** | 7 days | Days post-infection when detection occurs (`detect_time`) |
| **Recovery threshold** | 17 days | Infected individuals recover after this many days |
| **Initial S:I ratio** | 70:30 | Susceptible vs infected at start |

\newpage

## Model overview and key settings

This document implements an individual-based microsimulation (ABM) of an infectious disease with quarantine. The model tracks individuals through five health states over daily cycles.

- **Population**: `n_i = 1000` individuals, followed for `n_cycles = 60` daily cycles.
- **Health states** (5 total):
  - `S` = Susceptible (not infected, not in quarantine)
  - `SQ` = Susceptible in quarantine
  - `I` = Infected (not in quarantine)
  - `IQ` = Infected in quarantine
  - `R` = Recovered (absorbing, no further infection risk)
- **Contact structure**:
  - Each individual belongs to a **family** (`fid`) and, if working, a **workplace** (`wid`).
  - On each day, we first simulate **workplace exposure** (`work_exposure()`), then **household exposure** (`home_exposure()`).
  - For susceptibles (`S`), infection risk combines **work contacts** and **home contacts**.
  - For quarantined susceptibles (`SQ`), infection risk comes **only from home**, because they do not attend work.
- **Daily update logic** (per cycle):
  1. Reset exposure counters and simulate work contacts → update `expl`.
  2. Simulate household exposure given current infected family members → update `expl_home`.
  3. Compute individual transition probabilities with `Probs()` and sample new health states.
  4. Apply `Detection()` to move detected infected individuals into quarantine states.
  5. Apply `Quarantine_Check()` to start, continue, or end quarantine based on `qtime`.
  6. Update infection duration `v_Ts` and move long-infected individuals to `R`.

---

-

## Code update 2026-02-18

**Documentation & setup**
- Header: added model statistics table; setup chunk clears workspace when knitting (keeps env when running interactively)
- Packages: trimmed to dplyr, ggplot2, reshape2, knitr, darthtools

**Parameters & population**
- Section 03: removed unused `cycle_length`, `p_detect_overall`, `p_detect_inf`; single household transmission `p_transhh <- 0.05`
- `create_abmpop`: removed `Sex`, `Age_group`, `work_contact_*`, `home_contact_*`, `contact_type_w/h`; added `v_Ts`, `detect_time`, `new_detection`; initial infected get `v_Ts` sampled from 1 to `(detect_time - 1)`

**Exposure & transitions**
- `work_exposure`: excludes SQ and IQ from workplace mixing; assigns `expl` by `ptid` ,avoids misalignment.
- `home_exposure`: per-family infected count → `expl_home` for S/SQ; unchanged logic
- `Probs`: rows = individuals, cols = states; removed unused `v_Ts` argument
- `Detection`: resets `new_detection` each cycle; moves detected (v_Ts == detect_time) to quarantine (I→IQ, S→SQ)

---

## Code update 2026-02-11

- **Workplace contacts**: In `create_abmpop()`, each worker gets `contact_num` sampled from `0` to `(workplace_size - 1)`
- **Household exposure**: `home_exposure()` counts infected members per family; sets `expl_home` for susceptibles (S, SQ) in that family
- **Transition matrix**: In `Probs()`, `m_p_t` has rows = individuals, columns = states; each row sums to 1



---

# 01 Load packages

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE, eval = TRUE)
options(scipen = 999)   # Disable scientific notation for numeric output
if (!interactive()) rm(list = ls())   # Clear workspace when knitting (preserves env when running interactively)
```

```{r}
# Core packages: data manipulation, plotting, document generation
if (!require('pacman')) install.packages('pacman'); library(pacman)
p_load("dplyr", "ggplot2", "reshape2", "knitr", "pacman")

# DARTH tools: samplev(), check_transition_probability(), check_sum_of_transition_array()
p_load_gh("DARTH-git/darthtools")
```

# 02 Load functions

```{r}
```

# 03 Model input

This section defines all parameters used by the simulation: population size and horizon, time constants (quarantine, latent period), health state labels, state groupings for logic, and transmission probabilities.

## 03.1 Define model input parameters

```{r}
## General setup
set.seed(1)
n_cycles          <- 60                            # simulation horizon (daily cycles)
n_i               <- 1000                          # population size (individuals)

## Time parameters (days)
q_maxtime         <- 14                            # quarantine duration

## Health states (5 total)
v_names_states    <- c("S",                        # Susceptible
                       "SQ",                       # Susceptible in quarantine
                       "I",                        # Infected
                       "IQ",                       # Infected in quarantine
                       "R")                        # Recovered
v_names_cycles    <- paste("cycle", 0:n_cycles)
n_states          <- length(v_names_states)

## Custom operator: not-in
`%!in%`           <- Negate(`%in%`)

## State groups (Probs, Detection, Quarantine_Check)
ending_states     <- c("R")
infect_states     <- c("I", "IQ")
infect_states_iso <- c("IQ")                       # infected in quarantine only
all_infect_states <- c(infect_states, infect_states_iso)
iso_states        <- c("SQ","IQ")
sus_states        <- c("S", "SQ")

## Transmission probabilities (per contact)
p_transhh       <- 0.05                            # household
p_TransmitWork  <- 0.03                            # workplace
```

# 04 Sample individual level characteristics

## 04.1 Static characteristics

**Per-agent fields:**

| Field | Description |
|-------|-------------|
| `ptid` | Individual ID |
| `state` | Health state (S, SQ, I, IQ, R) |
| `fid` | Family ID |
| `wid` | Workplace ID |
| `qtime` | Quarantine clock (0 = not in Q; 1..q_maxtime = days in Q) |
| `time` | Cumulative total quarantine days |
| `expl` | Daily workplace exposure count (reset each cycle) |
| `expl_home` | Daily household exposure count (reset each cycle) |
| `contact_type` | 0/1: had workplace exposure this cycle (set by work_exposure) |
| `atwork` | Whether individual works |
| `contact_num` | Daily work contacts sampled (workers only) |
| `v_Ts` | Days since infection (0 if not infected) |
| `detect_time` | Day post-infection when detection occurs |
| `new_detection` | Newly detected this cycle (reset each cycle) |

```{r}
create_abmpop <- function(npop){
  # Returns: population dataframe with one row per individual

  agents <- data.frame(
    ptid   = 1:npop,
    state  = "S",
    wid    = 0,
    fid    = 0,
    qtime  = 0,
    time   = 0,
    atwork = TRUE
  )

  ## Initial state: 70% S, 30% I
  n_S   <- round(0.7 * npop)
  n_I   <- npop - n_S
  agents$state <- sample(c(rep("S", n_S), rep("I", n_I)))

  ## Family assignment (sizes 1, 2, 3, 4)
  family_size    <- c(1, 2, 3, 4)
  p_family_size  <- family_size / sum(family_size)
  
  # create equal number of families with different family size, and assign family id to each individual based on family size
  fam_num = floor(npop * p_family_size / family_size)[1]
  fam_id  = c(rep(1:fam_num,each = family_size[1]),
              rep((fam_num + 1): (fam_num*2),each=family_size[2]),
              rep((fam_num*2 + 1):(fam_num*3),each=family_size[3]),
              rep((fam_num*3 + 1):(fam_num*4),each=family_size[4]))
  
  fam_id = c(fam_id,rep((max(fam_id) + 1),each = (npop - length(fam_id))))
  
  # Workplace assignment (sizes 5, 7, 8)
  work_size   <- c(5, 7, 8)
  p_work_size <- work_size / sum(work_size)
  work_num = floor(npop * p_work_size / work_size)[1]
  work_id  = c(rep(1:work_num,each = work_size[1]),
               rep((work_num + 1): (work_num*2),each=work_size[2]),
               rep((work_num*2 + 1):(work_num*3),each=work_size[3]))
  
  work_id = sample(c(work_id,rep((max(work_id) + 1),each = (npop - length(work_id)))))
  
  
  agents$fid <- fam_id
  agents$wid <- work_id

  ## Exposure and contact (reset/updated each cycle)
  agents$expl         <- 0
  agents$expl_home    <- 0

  ## Infection and detection
  detect_time_val      <- 7                            # day post-infection when detection occurs
  agents$detect_time   <- detect_time_val
  agents$new_detection <- 0                            # newly detected this cycle (reset each cycle)
  agents$v_Ts          <- 0                            # time since infection (0 if not infected)
  
  # For initially infected: sample v_Ts from 1 to (detect_time - 1)
  idx_inf              <- agents$state == "I"
  agents$v_Ts[idx_inf] <- sample(1:(detect_time_val - 1), size = sum(idx_inf), replace = TRUE)
  
  ## Workplace contacts: sample daily contact count (0 to workplace_size - 1)
  work_in_sample <- agents$ptid[agents$atwork]
  work_size_tbl <- table(agents[work_in_sample, ]$wid)
  work_ppl        <- agents[work_in_sample, ]
  work_info       <- data.frame(
    wid   = as.numeric(names(work_size_tbl)),
    wsize = as.numeric(work_size_tbl)
  )
  work_place_size <- merge(work_ppl, work_info, by = "wid")
  work_place_size <- work_place_size[order(work_place_size$ptid), ]

  work_place_size <- work_place_size %>%
    rowwise() %>%
    mutate(contact_num = sample(0:(wsize - 1), 1))

  agents$contact_num <- NA
  agents$contact_num[work_in_sample] <- work_place_size$contact_num
  agents$contact_type <- 1

  return(agents)
}

```


## 04.2 Create a dataframe with the individual characteristics

```{r}
# Create a dataframe with the individual characteristics (e.g., family id, workplace id, initial health state, time of being sick at the start of the model) for every individual in the population.
df_X <- create_abmpop(n_i)

```

# 05 Define Simulation Functions

## 05.1 Probability function

The `Probs` function updates the transition probabilities of every cycle is shown below.

```{r}

infect_prob <- function(prob,exposure) { 
  # Arguments:
  # prob:     the probability of infection in one time contact
  # exposure: the time of being exposed to infectious individuals
  # Returns: 
  # the probability of being infected based on certain exposure time
  return(1 - (1 - prob) ^ (exposure))   # 1 - prob is the prob of being safe, 
  # 1 minus the prob of always being safe is the prob of getting infection
}

Probs <- function(M_t, df_X) { 
  # Arguments:
  # M_t:  health state occupied at cycle t (character variable)
  # df_X: data frame with individual characteristics data 
  # Returns: 
  # transition probabilities for that cycle
  
  # create matrix of state transition probabilities
  # rows = individuals, columns = states (this matches samplev() expectations)
  m_p_t           <- matrix(0, nrow = n_i, ncol = n_states)
  colnames(m_p_t) <- v_names_states                            
  
  #Calculate the probability of being infected for susceptible people
  sus <- df_X[df_X$state == "S",] #identify those susceptible
  
  #check exposure number
  sus_prob_home  <- rep(0, nrow(sus)) 
  contact_active <- sus[sus$expl > 0 | sus$expl_home > 0, ]
  at_home        <- contact_active$expl_home > 0
  
  if(sum(at_home) > 0){
    home_infect = infect_prob(p_transhh, contact_active[at_home,]$expl_home)
    sus_prob_home[sus$expl_home > 0] = home_infect
  }
  
  sus_prob = rep(0,nrow(sus))
  at_work = contact_active$contact_type == 1
  if(sum(at_work) > 0){
    work_infect = infect_prob(p_TransmitWork,contact_active[at_work,]$expl)
    sus_prob[sus$expl > 0 & sus$contact_type == 1] = work_infect
  }
  p_Sus_Inf <- (1- ((1- sus_prob_home) * (1 - sus_prob))) # 1 minus the prob of being safe all day
  

  #Calculate the probability of being infected for susceptible quarantine
  sus_Q <- df_X[df_X$state == "SQ",] #identify those susceptible
  
  #check exposure number
  sus_prob_home_Q  <- rep(0, nrow(sus_Q)) 
  contact_active_Q <- sus_Q[sus_Q$expl > 0 | sus_Q$expl_home > 0, ]
  at_home_Q        <- contact_active_Q$expl_home > 0
  
  if(sum(at_home_Q) > 0){
    home_infect_Q = infect_prob(p_transhh, contact_active_Q[at_home_Q,]$expl_home)
    sus_prob_home_Q[sus_Q$expl_home > 0] = home_infect_Q
  }

  
  
  
  # the 5 health states of the model:
v_names_states  <- c("S",   # Susceptible 
                     "SQ",  # Susceptible quarantine 
                     "I",   # Infected
                     "IQ",  # Infected quarantine
                     "R")   # Recovered      

  
  #susceptible 
  m_p_t[M_t == "S", "S" ]                     <- 1 - p_Sus_Inf
  m_p_t[M_t == "S", "SQ"]                     <- 0
  m_p_t[M_t == "S", "I" ]                     <- p_Sus_Inf
  m_p_t[M_t == "S", "IQ"]                     <- 0
  m_p_t[M_t == "S", "R" ]                     <- 0
  
  
  #susceptible quarantine
  m_p_t[M_t == "SQ", "S" ]                    <- 0
  m_p_t[M_t == "SQ", "SQ"]                    <- 1 -  sus_prob_home_Q
  m_p_t[M_t == "SQ", "I" ]                    <- 0
  m_p_t[M_t == "SQ", "IQ"]                    <- sus_prob_home_Q
  m_p_t[M_t == "SQ", "R" ]                    <- 0
  
  #infected
  m_p_t[M_t == "I" , "S" ]                    <- 0
  m_p_t[M_t == "I" , "SQ"]                    <- 0
  m_p_t[M_t == "I" , "I" ]                    <- 1
  m_p_t[M_t == "I" , "IQ"]                    <- 0
  m_p_t[M_t == "I" , "R" ]                    <- 0
  
  #infected quarantine
  m_p_t[M_t == "IQ", "S" ]                    <- 0
  m_p_t[M_t == "IQ", "SQ"]                    <- 0
  m_p_t[M_t == "IQ", "I" ]                    <- 0
  m_p_t[M_t == "IQ", "IQ"]                    <- 1
  m_p_t[M_t == "IQ", "R" ]                    <- 0
  
  
  #recover ending state
  m_p_t[M_t == "R" , "S" ]                    <- 0
  m_p_t[M_t == "R" , "SQ"]                    <- 0
  m_p_t[M_t == "R" , "I" ]                    <- 0
  m_p_t[M_t == "R" , "IQ"]                    <- 0
  m_p_t[M_t == "R" , "R" ]                    <- 1
   
  
  # Return transition probabilities with rows = individuals, cols = states
  return(m_p_t)
}  


```

## 05.2 Quarantine Function

```{r}

#Quarantine_Check can update the quarantine clock at current time cycle
Quarantine_Check <- function(pop, pre_state){
  # Arguments:
  # pop: population dataframe of current time cycle t
  # pre_state: states at previously cycle t-1
  # Returns: 
  # new population dataframe with updated states
  
  # Step1: find people in different quarantine conditions
  
  # People in quarantine:
  # condition 1: reset quarantine clock for those who become infected or have close contact with infected people during quarantine
  new_infected_in_Q          = which(pop$state %in% infect_states_iso & pre_state == "SQ") #got infected in quarantine
  
  reset_family               = unique(pop$fid[new_infected_in_Q])
  reset_people               = pop$fid %in% reset_family #find all the people who had expose in family quarantine
  
  
  # condition 2: find people are going to be released from quarantine
  # check if quarantine clock hit q_maxtime
  full_q                     = pop$qtime == q_maxtime & !reset_people & pop$state %!in% ending_states 
  
  #find susceptible people and infected people who meet quarantine time
  sus_full_q                 = full_q & pop$state == "SQ"
  inf_full_q                 = full_q & pop$state %in% infect_states_iso 
  
  # condition 3: find people who stay in quarantine and the quarantine clock has to be incremented.
  else_q                     = (pop$qtime < q_maxtime & pop$qtime >0) & !reset_people & pop$state %!in% ending_states  
  
  # Other people:
  # condition 4: find newly infected people entering quarantine NOTE: people enter quarantine by detection
  new_infected               = which(pop$state %in% iso_states & pop$qtime == 0) #new infected generated from infected
  
  # condition q_maxtime: find newly infected people's dependencies in families and workplaces
  new_infected_family        = unique(pop$fid[new_infected]) #families have newly infected in quarantine cases
  
  #new_infected_work= unique(pop$wid[new_infected]) #workplaces have newly infected in quarantine cases #comment out if same workplace contact don't need to be put in quarantine
  
  new_qurantine              = (((pop$fid %in% new_infected_family & pop$qtime == 0) | #(pop$wid %in% new_infected_work & pop$qtime == 0) |
                                   (pop$fid %in% reset_family)) & (pop$state %!in% ending_states)) #people whose quarantine clocks have to be set to 1 including
  new_qurantine_add = new_qurantine & pop$state %!in% iso_states #find people need to be put in quarantine state
  
  
  # Step2: Update quarantine clock
  # set new_quarantine clock to 1 
  pop$qtime[new_qurantine]    = 1
  pop$state[new_qurantine_add] = paste(pop$state[new_qurantine_add],"Q",sep = "") #update quarantine states
  pop$time[new_qurantine]     = pop$time[new_qurantine]+1 #increment total quarantine time
  
  
  # release people with meet q_maxtime-days quarantine requirement
  # - isolated susceptible people will go to susceptible
  # - infected people will go to recovered
  pop$qtime[full_q]           = 0
  pop$state[sus_full_q]       = "S" #when quarantine ends, state goes back to susceptible
  pop$state[inf_full_q]       = "R" #when quarantine ends, state goes back to recovered
  
  #increment quarantine clock and total quarantine time for everyone else in quarantine
  pop$qtime[else_q]           = pop$qtime[else_q] + 1
  pop$time[else_q]            = pop$time[else_q] + 1
  return (pop)
}

```

## 05.3 Exposure at work

```{r}

#Exposure_Check Work
work_exposure <- function(pop){
  allpop                      = pop
  
  if(sum(pop$atwork == 1) > 0){
    select_pop = pop[pop$atwork & pop$state %!in% c("SQ", "IQ"), ]
    expl_list                 = select_pop$expl
    infectious                = select_pop$state %in% infect_states # defn of infectious
    wid_list                  = select_pop$wid
    out_of_iso                = select_pop$state %!in% iso_states
    ptid_list                 = select_pop$ptid[out_of_iso & select_pop$state == "S"]
    atwork                    = select_pop$atwork
    
    workplacetable                = data.frame(table(wid_list[atwork]))
    
    names(workplacetable)         = c("wid","wsize")
    
    inf_worker               = select_pop[infectious ,]
    inf_workplace                = inf_worker$wid
    
    if (nrow(inf_worker) > 0){
      contact                   <- c()
      
      for( i in c(1:nrow(inf_worker))){
        select_worker           = inf_worker[i,]
        workplaceptid           = select_pop[select_pop$wid == select_worker$wid,]$ptid
        to_contact              = workplaceptid[workplaceptid %!in% select_worker$ptid]
        if (length(to_contact) > 1){
          new_contact = sample(to_contact,
                               min(length(to_contact),select_worker$contact_num))
          contact                 = c(contact,new_contact)
        }else if (length(to_contact) == 1){
          new_contact = to_contact
          contact                 = c(contact,new_contact)
        }
      }
      
      if (length(contact) != 0){
        contact_table             <- data.frame(table(contact))
        names(contact_table)      = c("ptid","expl")
        
        contact_table$ptid        <- as.numeric(as.character(contact_table$ptid))
        contact_table$expl        <- as.numeric(as.character(contact_table$expl))
        
        expl_list[select_pop$ptid %in% contact_table$ptid] = expl_list[select_pop$ptid %in% contact_table$ptid] + contact_table$expl
        allpop[(allpop$ptid %in% contact_table$ptid ),]$contact_type = 1
      }
    
    }
    allpop$expl[allpop$ptid %in% select_pop$ptid] <- expl_list
    
  }
  
  return(allpop) 
}
```

## 05.4 Exposure at home

```{r}
home_exposure <- function(pop){
  # Arguments:
  # pop: population dataframe of current time cycle t
  # Returns:
  # population dataframe with updated home exposure level (`expl_home`)
  #
  # Logic:
  # - For each family (fid), count how many infected members are present
  #   (states in `all_infect_states`, e.g. I or IQ).
  # - For each susceptible or susceptible-quarantine individual (S, SQ) in
  #   that family, set `expl_home` equal to the number of infected family
  #   members on that day.
  
  # reset home exposure
  pop$expl_home <- 0
  
  # identify infected individuals
  inf_idx <- pop$state %in% all_infect_states
  
  if (any(inf_idx)) {
    # count infected members per family
    fam_counts <- table(pop$fid[inf_idx])
    fam_ids    <- as.numeric(names(fam_counts))
    
    for (i in seq_along(fam_ids)) {
      fid_i <- fam_ids[i]
      n_inf <- as.numeric(fam_counts[i])
      in_family_sus <- pop$fid == fid_i & pop$state %in% c("S", "SQ")
      pop$expl_home[in_family_sus] <- n_inf
    }
  }
  
  return(pop)
}

```

## 05.5 Detection function

The `Detection` function identifies infected individuals who are detected (e.g., via testing) on the current day and moves them into quarantine.

```{r}
Detection <- function(pop, v_Ts){
  # pop  : population at time t (after updating infection status)
  # v_Ts : days since infection for each individual
  
  # Reset per-cycle detection counter
  pop$new_detection <- 0
  
  # Time since infection equals detection time and not in an ending state → detected
  detection_period    <- v_Ts 
  detected_individuals <- (detection_period == pop$detect_time) &
                          pop$state %!in% ending_states

  # Count newly detected and move non-isolated detected individuals into quarantine
  pop$new_detection   <- detected_individuals
  new_iso             <- detected_individuals & pop$state %!in% iso_states
  pop$state[new_iso]  <- paste(pop$state[new_iso], "Q", sep = "")

  return(pop)
}

```

## 05.6 Microsimulation function

Below we develop the microsimulation function that allows the model to be run.

```{r}
MicroSim <- function(n_i = 100, df_X, seed = 1, n_cycles = 10) {
  # Arguments:  
    # n_i: number of individuals
    # df_X: data frame with individual data 
    # seed: seed for the random number generator, default is 1
    # n_cycles: number of cycles to simulate
  # Returns:
    # results: list with m_M (state trace), pop, v_new_inf_daily, v_total_inf_daily, v_detected_daily
  
  set.seed(seed) # set a seed to be able to reproduce the same results
  
  # m_M stores health state over time for every individual (rows = individuals, cols = cycles)
  # m_vts stores time since infection; m_expl stores work exposure per cycle
  
  m_M <-  matrix(nrow = n_i, ncol = n_cycles + 1, 
                               dimnames = list(paste("ind"  , 1:n_i, sep = " "), 
                                               paste("cycle", 0:n_cycles, sep = " ")))  
 
  m_vts    <-  matrix(nrow = n_i, ncol = n_cycles + 1, #tracking number infected days
                    dimnames = list(paste("ind"  , 1:n_i, sep = " "), 
                                    paste("cycle", 0:n_cycles, sep = " ")))   
  
  m_expl <-  matrix(nrow = n_i, ncol = n_cycles, 
                 dimnames = list(paste("ind"  , 1:n_i, sep = " "), 
                                 paste("cycle", 1:n_cycles, sep = " ")))  

  m_M[, 1] <- as.character(df_X$state)    # initial health state at cycle 0 for individual i
  m_vts[,1]        <- df_X$v_Ts  
  v_Ts             <- df_X$v_Ts           # initialize time since illness onset

  v_total_inf_daily <- rep(0,n_cycles)    # initialize the number of total infections per cycle
  v_new_inf_daily   <- rep(0,n_cycles)    # initialize the number of total new infections per day
  v_detected_daily  <- rep(0,n_cycles)    # initialize the number of total infections per cycle
  
  
  # open a loop for time running cycles 1 to n_cycles 
  for (t in 1:n_cycles) {
     print(paste("running day: ",as.character(t),sep = ""))
    
    #check exposure
    df_X$expl            <- 0                                 # reset expl level
    df_X$expl_home       <- 0                                 # reset expl level
    df_X$contact_type    <- 0                                 # reset contact type
    df_X <- work_exposure(df_X)
    df_X <- home_exposure(df_X)
    m_expl[,t] <- df_X$expl
    
    
    # calculate the transition probabilities for the cycle based on health state t
    m_P <- Probs(m_M[, t], df_X)
    # check if transition probabilities are between 0 and 1
    check_transition_probability(m_P, verbose = TRUE)
    # check if each of the rows of the transition probabilities matrix sum to one
    check_sum_of_transition_array(m_P, n_rows = n_i, n_cycles = n_cycles, verbose = TRUE)
    
    prob_new_state = samplev(m_P) #generate new states based on the prob
    
    df_X$state = prob_new_state
    df_X = Detection(df_X,v_Ts)
    
    prob_new_state = df_X$state
    
    v_detected_daily[t] = sum(df_X$new_detection)
    v_total_inf_daily[t] = sum(prob_new_state %in% all_infect_states)
    v_new_inf_daily[t] = sum((m_M[, t] %in% sus_states   & prob_new_state %in% all_infect_states ))
    
    
    #update Quarantine Status
    df_X = Quarantine_Check(df_X,m_M[, t]) #m_M[,t] is the previous day's state
    qc_state = df_X$state # state_after quarantine check
    
    # update time since illness onset for t + 1 
    v_Ts <- if_else(qc_state %in% all_infect_states, v_Ts + 1, 0)
    m_vts[,t + 1] =  v_Ts 
    
    v_go_to_recover <- if_else(v_Ts >= 17, TRUE, FALSE)
    df_X$state[v_go_to_recover] = "R"
    
    m_M[, t + 1]  <- df_X$state  # new day's state
   
    
    # Display simulation progress
    if(t/(n_cycles/10) == round(t/(n_cycles/10), 0)) { # display progress every 10%
      cat('\r', paste(t/n_cycles * 100, "% done", sep = " "))
    }
    
  } # close the loop for the time points 
  
  # store the results from the simulation in a list
  results <- list(m_M = m_M ,
                  pop = df_X, 
                  v_new_inf_daily = v_new_inf_daily, 
                  v_total_inf_daily =   v_total_inf_daily ,
                  v_detected_daily =  v_detected_daily)   
  
  return(results)  # return the results

} # end of the `MicroSim` function  
```

## 06 Run Microsimulation

```{r, eval = TRUE, echo = TRUE, warning = FALSE, message = FALSE, results = FALSE }
# 06 Run Microsimulation
outcomes <- MicroSim(n_i = n_i, df_X = df_X, seed = 1, n_cycles = n_cycles)

```

```{r, eval = TRUE, echo = TRUE, warning = FALSE, message = FALSE}
# 06.1 Plot state trends (S, I, Q)

state_mat <- outcomes$m_M
cycles <- 0:(ncol(state_mat) - 1)

count_state <- function(s) colSums(state_mat == s)

df_trend <- data.frame(
  cycle = cycles,
  S  = count_state("S"),
  I  = count_state("I"),
  SQ = count_state("SQ"),
  IQ = count_state("IQ"),
  R  = count_state("R")
)
df_trend$Q <- df_trend$SQ + df_trend$IQ

# Plot S, I, Q (quarantine = SQ + IQ)
df_long <- reshape2::melt(df_trend, id.vars = "cycle", measure.vars = c("S","I","Q"))
df_long$variable <- factor(df_long$variable, levels = c("S","I","Q"),
                           labels = c("Susceptible","Infected","Quarantine"))
ggplot(df_long, aes(x = cycle, y = value, color = variable)) +
  geom_line() +
  labs(title = "State Trends: Susceptible, Infected, Quarantine",
       x = "Cycle (day)",
       y = "Number of individuals",
       color = "State")

# plot all states separately
df_long_all <- reshape2::melt(df_trend, id.vars = "cycle", measure.vars = c("S","SQ","I","IQ","R"))
ggplot(df_long_all, aes(x = cycle, y = value, color = variable)) +
  geom_line() +
  labs(title = "State Trends: All States",
       x = "Cycle (day)",
       y = "Number of individuals",
       color = "State")
```

```{r, eval = TRUE, echo = TRUE, warning = FALSE, message = FALSE}
# 06.2 Daily counts summary table

df_daily <- data.frame(
  day = cycles,
  susceptible = df_trend$S,
  infected = df_trend$I,
  recovered = df_trend$R,
  infected_Q = df_trend$IQ,
  sus_Q = df_trend$SQ
)
df_daily
```

We kindly request you to add the following Acknowledgement paragraph to your further work where DARTH code formed the basis. We also like to remind you that you can add other sources of reference to this paragraph to acknowledge code you got from others.

# Acknowlegdement

For this work we made use of the template developed by the Decision Analysis in R for Technologies in Health (DARTH) workgroup: <http://darthworkgroup.com>.

The notation of our code is based on the following provided framework and coding convention: Alarid-Escudero, F., Krijkamp, E., Pechlivanoglou, P. et al. A Need for Change! A Coding Framework for Improving Transparency in Decision Modeling. PharmacoEconomics 37, 1329--1339 (2019). <https://doi.org/10.1007/s40273-019-00837-x>.

Other work from DARTH can be found on the website: <http://darthworkgroup.com/publications/>

# Copyright for assignment work

Copyright 2017, THE HOSPITAL FOR SICK CHILDREN AND THE COLLABORATING INSTITUTIONS.All rights reserved in Canada, the United States and worldwide. Copyright, trademarks, trade names and any and all associated intellectual property are exclusively owned by THE HOSPITAL FOR Sick CHILDREN and the collaborating institutions. These materials may be used, reproduced, modified, distributed and adapted with proper attribution.
